Primer on Production Functions and Economic Growth

### 2014-11-30 Articles/Prod Func Primer/ProdFuncPrimer.tex
Edited the Primer.

The Cobb-Douglas production function is given by
%
\begin{equation} \label{eq:CD}
  y = A k^\alpha l^\beta e^\gamma \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; .
\end{equation}
%
$A$ is known as total factor productivity,
and $\lambda$ is the Solow residual.

In the Cobb-Douglas world, factors of production are substitutable, 
eg., a reduction of labor~($l$) can be compensated by an increase of capital~($k$)
without affecting output~($y$).%
  \footnote{
  In fact, the capital-for-labor substitution has a name: automation.
  }
Furthermore, an increase in one factor of production (say, $k$)
without a concomitant increase in the others ($l$ or $e$) 
will result in an increase of output ($y$), provided that the output elasticity 
($\alpha$) is non-zero.

@@

%++++++++++++++++++++++++++++++
\subsection{CES with Two Factors of Production} 
\label{sec:CES-2}
%++++++++++++++++++++++++++++++

@@

%++++++++++++++++++++++++++++++
\subsection{CES with Three Factors of Production} 
\label{sec:CES-3}
%++++++++++++++++++++++++++++++

@@

@@

\begin{equation} \label{eq:LINEX_beta}
  \beta \equiv \frac{l}{y} \frac{\partial y}{\partial l} = a \left[ c \left( \frac{l}{e}\right) - \frac{l}{k} \right] \;,
\end{equation}
%
\noindent{}and
%
\begin{equation} \label{eq:LINEX_beta_2}
  \gamma \equiv \frac{e}{y} \frac{\partial y}{\partial e} 
         = 1 - a \left[ \frac{e}{k} + c \frac{l}{e} \right] 
         = 1 - \alpha - \beta \;.
\end{equation}
%
The LinEx model assumes constant returns to scale.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Appendix}

Include data tables here?

>>>>

The Cobb-Douglas production function is given by
%
\begin{equation} \label{eq:CD}
  y = A k^\alpha l^\beta e^\gamma \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; .
\end{equation}
%
$A$ is known as total factor productivity,
and $\lambda$ is the Solow residual.
Constant returns to scale is represented by $\alpha + \beta + \gamma = 1$.

%++++++++++++++++++++++++++++++
\subsection{Cobb-Douglas and substutitability} 
\label{sec:CD-substitution}
%++++++++++++++++++++++++++++++

In the Cobb-Douglas world, factors of production are substitutable, 
eg., a reduction of labor~($l$) can be compensated by an increase of capital~($k$)
without affecting output~($y$).%
  \footnote{
  In fact, the capital-for-labor substitution has a name: automation.
  }
Furthermore, an increase in one factor of production (say, $k$)
without a concomitant increase in the others ($l$ or $e$) 
will always result in an increase of output ($y$), provided that the output elasticity 
($\alpha$) is non-zero.

%++++++++++++++++++++++++++++++
\subsection{The cost share theorem} 
\label{sec:CD-cost-share}
%++++++++++++++++++++++++++++++

The so-called ``cost share theorem'' says that coefficients ($\alpha$, $\beta$, and $\gamma$) 
in the Cobb-Douglas model are equal to cost shares in the economy.~\citep[p.~150]{Kummel:2010vz}
\citet[Eqn.~12]{Kummel:2010vz} shows (through the use of Lagrange multipliers) 
that the ``cost share theorem'' is strictly true for a profit maximizing economy
if and only if there are \emph{no technological constraints} on economic production.

There are many technological constraints that affect economic production, such as
the Second Law of Thermodynamics and
the productive capability and energy efficiency 
of capital~($k$) at any instant in time.

@@

%++++++++++++++++++++++++++++++
\subsection{CES with two factors of production} 
\label{sec:CES-2}
%++++++++++++++++++++++++++++++

@@

%++++++++++++++++++++++++++++++
\subsection{CES with three factors of production} 
\label{sec:CES-3}
%++++++++++++++++++++++++++++++

@@

%++++++++++++++++++++++++++++++
\subsection{CES assumptions} 
\label{sec:CES_assumptions}
%++++++++++++++++++++++++++++++

The CES production function was derived by assuming profit maximization
and perfect competition in the markets for both output and 
factors of production.~\citep[p.~228]{Arrow:1961aa}
In particular the assumption that 
``the real wage equals the marginal product of labor''~\citep[p.~6]{Klump:2011aa}
is made.

As I have written before, these assumptions become problematic
when considering energy.
Many countries have parastatal or full government control 
over electricity generation and/or fuel markets.  
Many countries subsidize energy markets.  
And, there are always health and environmental externalities 
that are not priced into fossil fuel energy markets.  
So, the conditions for perfect competition are not present for energy markets.

@@

\begin{equation} \label{eq:LINEX_beta}
  \beta \equiv \frac{l}{y} \frac{\partial y}{\partial l} 
  = a \left[ c \left( \frac{l}{e}\right) - \frac{l}{k} \right] \;,
\end{equation}
%
and
%
\begin{equation} \label{eq:LINEX_beta_2}
  \gamma \equiv \frac{e}{y} \frac{\partial y}{\partial e} 
         = 1 - a \left[ \frac{e}{k} + c \frac{l}{e} \right] 
         = 1 - \alpha - \beta \;.
\end{equation}
%
The LinEx model assumes constant returns to scale.

@@

Common forms of macroeconomic production functions have been reviewed.
Relationships among production functions have been examined.
Key assumptions for each have been discussed.

@@

After reviewing this literature, I think there is a need
for a book entitled \emph{Energy in the Production Function}
which would start with Equation~\ref{eq:diff_growth}
and carefully derive each of the production functions.
In each derivation, assumptions should be clearly exposed.
The constraints associated with the LiNex equation should be critically
examined.
New constraints should be introduced to round out the picture of 
economic growth in an energy- and materials-constrained world.
If possible, the price elasticity of supply, especially for energy, should be 
included in the production function.
I'm unsure how this should be done, 
although a profitable route might be found by
re-thinking the marginal product of energy.
Finally, a new form of the production function should be derived
and applied.

This new piece should be an entry point for people interested 
in the role of energy in macroeconomic growth
who have a differential equations background.
\emph{Lecture Notes in Energy} might be a good home for it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Acknowledgements}
% \label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \section{Appendix}
%
% Include data tables here?
